---
title: 'Tutorial 1: Introduction to Machine Learning with Python¶'
author: "Jonathan Ish-Horowicz"
date: "12/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tutorial 1: Introduction to Machine Learning with Python

The goal of this tutorial is to introduce a typical workflow in carrying out ML in R.  Similarly to last week's Python tutorial, this includes:

1. accessing and organising data,
2. assessing the data,
3. visualising the data,
4.  a) creating training, b) test datasets and c) learning a model using them and evaluating its performance.

# 1) Load Data

We load the same iris dataset as last week, which has 150 samples and 4 attributes. There are 3 classes (species).

In R we can load the Iris dataset from the `datasets` package:

```{r message=FALSE}
# Load the iris dataset
library(datasets)
data(iris)
iris$Species <- as.character(iris$Species)
```

# 2) Statistics of the dataset

Now compute the mean, standard deviation, minimum and maximum of each attribute.

Suggestion: use the the `group_by` and `summarise_all` functions from the `dplyr` library.

```{r message=FALSE}
library(dplyr)

# Calculate the mean of each attribute
iris %>% group_by(Species)%>% 
  summarise_if(is.numeric, mean)

# Calculate the standard deviation of each attribute
iris %>% group_by(Species) %>% 
  summarise_if(is.numeric, sd)

# Calculate the minimum of each attribute
iris %>% group_by(Species) %>% 
  summarise_if(is.numeric, min)

# Calculate the maximum of each attribute
iris %>% group_by(Species) %>% 
  summarise_if(is.numeric, max)
```
Alternative input into the summarise function is summarise_all(list(mean))


# 3) Visualise the dataset

Make some exploratory plots here.

Suggestion: use the `ggplot2` library. For a nice pairs plot use the `ggpairs` function from `GGally` (a `ggplot2` extension library).

```{r EDA, message=FALSE}
library(ggplot2)
library(GGally)
ggpairs(iris, aes(color = Species, alpha = 0.5), progress=FALSE)
  #alpha = 0.5 needed so you can see the multiple plots
```
Plot shows that it should be relatively easy to classify, as the dots group themselves quite clearly.

# 4) Classification using Least Squares

Here we will be carrying out classification using the least squares formulation on 2 classes of the dataset.

a) Create separate datasets for the classes `setosa` and `versicolor`.

```{r}
# your code here

# result should be two dataframes (one for each for setosa, versicolor classes),
# each with dim (50,5) - 4 attributes plus column for class
#stopifnot(all(dim(setosa)==c(50,5)))
#stopifnot(all(dim(versicolor)==c(50,5)))
```
Don't need some of the code here as have done the ifelse statement

b) add a column to each dataset where the column is $1$ if the class is `setosa` and $-1$ otherwise.

```{r}
# your code here
iris$Class <- ifelse(iris$Species == 'setosa', 1, -1)
irissub <- subset(iris, Species != 'virginica')
# result should add a column to each of setosa and versicolor
#stopifnot(all(dim(setosa)==c(50,6))) #Test
#stopifnot(all(dim(versicolor)==c(50,6))) #Test
```


c) create training and test datasets, with 20% of the data for testing. This 80 training points and 20 testing points in total (half this per class).

```{r}
# your code here
set.seed(123) # Set Seed so that same sample can be reproduced in future also
library(caTools)
# Now Selecting 75% of data as sample from total 'n' rows of the data  

sample <- sample.int(n = nrow(irissub), size = floor(.80*nrow(irissub)), replace = F)
train <- irissub[sample, ]
test  <- irissub[-sample, ]
# resulting dataframes (one each for training and test data) should have
# the appropriate sizes
#stopifnot(all(dim(training.data)==c(80,6)))
#stopifnot(all(dim(test.data)==c(20,6)))
```

d) apply the least squares solution to obtain an optimal solution for different combinations of the 4 available attributes. The code to create a list containing all the combinations of the attributes has been provided.

```{r}
# Creates all possible combinations of attributes
# attribute.combinations is a list whose elements are lists of attributes
attribute.names <- colnames(iris)[1:4]
attribute.combinations <- do.call(
  c,
  lapply(1:4, function(i) as.list(data.frame(combn(attribute.names, i))))
  )
names(attribute.combinations) <- 1:length(attribute.combinations)
```
This created a list of the 4 attribute names, and then combined them in all possible combinations. 

Want to do:
(transposed train x train)inversed x (transposed train x class)
```{r}
return.predictions <- function(attribute.names, training.data, test.data) {
  
  # Format training and test data (as matrices)
  ### your code here
  trainmat <- data.matrix(train[-5])
  trainx <- trainmat[,1:4]
  trainy <- trainmat[,5]
  testmat <- data.matrix(test[-5])
  testy <- testmat[,1:4]
  testy <- testmat[,5]
  # Calculate optimal weights
  ### your code here
  #classifier <- glm(log(iris$Class) ~ iris$Petal.Length*iris$Petal.Width, family="binomial", data = iris)
  weights <- solve((t(trainx)%*%trainx)) %*% (t(trainx)%*%trainy)
  weightsmat <- data.matrix(weights)
  # Make predictions
  ###
  predictions <-  trainx %*% weights
  
    return(predictions)
}
```
### Notes
Classifier algorithm did not work. Trying matrix multiplication instead
Needs to be a matrix, different to a dataframe in R. 
Lecturer recommends not using a for loop - much faster when have lots of data.  Didn't do what we did, instead  used sapply on attribute combinations. Can then plot all of the combinations against their weights in a bar plot.  
Need to compare to the answers in teh complete data. 

e) evaluate which input attributes are the best.

```{r}
# Calculate the mean square error between some predictions and
# the corresponding testing data
return.mse <- function(predictions, testing.data) {
  ### your code 
  rmse(testmaty, predictions)
  
  return(mse)
}

# Calculate the test MSE for each the elements of attribute.combinations
# by calling return.predictions and return.mse



```